### 2013-10-12 research/paper/figure/m_star_collapsed.tex
uploaded 12-10-2013

\begin{figure}

\setlength{\unitlength}{\textwidth}
\fbox{
  \begin{picture}(1,0.32)

       \put(0.07,0.03){\includegraphics[width=0.3\unitlength]{../FnP/gnuplot/mean_power_collapsed_mstar.eps}}
       \put(0.36,0.03){\includegraphics[width=0.3\unitlength]{../FnP/gnuplot/mean_power_collapsed_noshed_mstar.eps}}
       \put(0.65,0.03){\includegraphics[width=0.3\unitlength]{../FnP/gnuplot/fsi_power.eps}}

        \put(0.03,0.17){\large $\frac{P_{m}}{\rho \mathcal{A}U^3}$}

        \put(0.2,0.0){$\frac{c}{\rho\mathcal{A}U}$} 	
        \put(0.49,0.0){$\frac{c}{\rho\mathcal{A}U}$}
        \put(0.78,0.0){$\frac{c}{\rho\mathcal{A}U}$}

        \put(0.133,0.21){\small(a)}
        \put(0.42,0.21){\small(b)}
        \put(0.715,0.21){\small(c)}

  \end{picture}
  }

  \caption{Mean power as a function of damping factor. Data are presented at $m^*=10$ (\ding{108}), $m^*=20$ (\ding{83}), $m^*=40$ (\ding{115}), $m^*=60$ (+) at Re 165 and $\zeta=0.1$. A reduction of maximum mean power can be observed when $m^*<40$. For $m^*>40$, the maximum power is essentially independent of $m^*$.}
    \label{fig:m_star_collapsed}
\end{figure}

Mean power as a function of damping factor at Re