### 2013-10-22 research/paper/figure/mstar_parkinson.tex

\begin{figure}
  \setlength{\unitlength}{\textwidth}
\fbox{
        \begin{picture}(1,0.3)(0.0,0.45)

      % % % Parkinson Data 
      \put(0.025,0.5){\includegraphics[width=0.5\unitlength]{../FnP/gnuplot/mean_power_collapsed_mstar_175.eps}}      \put(0.495,0.5){\includegraphics[width=0.5\unitlength]{../FnP/gnuplot/mean_power_collapsed_parkinson_10.eps}}

      \put(0.23,0.48){ $\displaystyle\frac{c}{\rho\mathcal{A}U}$}
      \put(0.73,0.48){ $\displaystyle\frac{c}{\rho\mathcal{A}U}$}

      \put(0.0,0.63){\large$\frac{P_{m}}{\rho \mathcal{A}U^3 }$}

      \put(0.085,0.709){\small(a)}
      \put(0.555,0.709){\small(b)}

    \end{picture}
}
  \caption{}
    \label{fig:collpased_data}
\end{figure}

\ %vspace{10cm}