### 2013-10-31 research/paper/figure/m_star_collapsed.tex
uploaded 31 OCt 2013

\begin{figure}

\setlength{\unitlength}{\textwidth}

  \begin{picture}(1,0.3)(0.0,0.45)

       \put(0.025,0.5){\includegraphics[width=0.5\unitlength]{../FnP/gnuplot/mean_power_collapsed_mstar.eps}}
       \put(0.495,0.5){\includegraphics[width=0.5\unitlength]{../FnP/gnuplot/mean_power_collapsed_noshed_mstar.eps}}

        \put(-0.01,0.637){\large $\frac{P_{m}}{\rho \mathcal{A}U^3}$}

        \put(0.23,0.48){$\displaystyle\frac{c}{\rho\mathcal{A}U}$} 	
        \put(0.73,0.48){$\displaystyle\frac{c}{\rho\mathcal{A}U}$}

     \put(0.085,0.709){\small(a)}
     \put(0.555,0.709){\small(b)}

  \end{picture}

  \caption{Mean power as a function of damping factor. Data are presented at $m^*=10$ (\ding{108}), $m^*=20$ (\ding{83}), $m^*=40$ (\ding{115}), $m^*=60$ (+) at Re 165 (a) with and (b) without the shedding term in Eq.\ref{final_equation_motion}. A reduction of maximum mean power can be observed when $m^*<40$ with shedding while the maximum power is essentially independent of $m^*$ when shedding is disregarded.}
    \label{fig:m_star_collapsed}
\end{figure}