### 2013-11-12 research/paper/results/results-2.3.tex

\section{Results}

\subsection{Stationary data}

The stationary $C_y$ data obtained using flow simulations (Fig.\ref{cy ploynomial}(a)) have a good agreement with data presented in \cite{Joly2012}. In comparison with the 7th order polynomial curve at \reynoldsnumber=22300 (Fig.\ref{cy ploynomial}(b)),  several differences could be observed. The peak value of $C_y$ is  significantly low at \reynoldsnumber=165 ($C_y=0.05$ at $4^0$) in comparison with \reynoldsnumber=22300 ($C_y=0.57$ at $13^0$) . The inflection point present at \reynoldsnumber=22300 could not be observed at \reynoldsnumber=165. This agrees with the findings of \cite{Luo2003}. It was concluded by \cite{Luo2003} that hysteresis occur due to the inflection point found in the $C_y$ curve. Therefore hysteresis is not expected at \reynoldsnumber=165. The range of the incident flow angles where $C_y$ remain positive is narrow at \reynoldsnumber=165 ($0^0 <\theta \leq$ $6^0$) compared to \reynoldsnumber=22300 ($0^0 <\theta \leq 15^0$). This feature is what sustains galloping and power is transferred from the fluid to the supporting structure within this range of incident angles because fluid forces are acting in the direction of travel of the oscillating cylinder. Incident angles beyond this range actually suppresses the galloping and power goes in the opposite direction. Therefore due to the overall smaller $C_y$ and narrow range of angles where $C_y$ is positive for \reynoldsnumber=165 compared to \reynoldsnumber=22300 leads one to expect a significant reduction in power at\reynoldsnumber=165 compared to \reynoldsnumber=2230.

\input{../figure/uncollapsed_data}

\subsection{Displacement,velocity and power output as a function of reduced velocity}

 The quasi-steady analysis data reveals that the displacement amplitude tend to grow with increasing $U^*$ (Fig.\ref{fig:uncollapsed_data} (a) and (b)). The onset of galloping is delayed with increasing $\zeta$ for both high and low Reynolds numbers. This echo the findings of previous studies by \cite{Parkinson1964} and \cite{Barrero-Gil2010a}. Hysteresis could be observed for the case with a higher Reynolds number. This was achieved by manipulating the initial conditions (initial displacement) of the system. The upper branch was obtained by giving an initial displacement which was higher than the expected amplitude while the lower branch was obtained by providing a lower initial displacement than the expected amplitude. Although a third state could be achieved theoretically, it was not possible to achieve numerically. This was also observed by \cite{Vio2007}.

 \subsubsection*{Power vs $U^*$}

 The mean power grows, peaks and then reduces as $U^*$ is increased (Fig.\ref{fig:uncollapsed_data} (e) and (f)) for each value of $\zeta$. The value of  \ustar at which the peak power occur increases with $\zeta$. However, the magnitude of the peaks remain constant for all the values of $\zeta$.  \cite{Barrero-Gil2010a} also showed a similar trend. The higher Reynolds number case clearly showed hysteresis in the power data. The range of hysteresis tend to increase with increasing $\zeta$. It could be observed that unlike VIV the  system has no preferred frequency. Although the onset of galloping and the point where peak power occur at different $U^*$ when the damping ratio is changed, the peak power remains constant regardless of $U^*$.

 \subsection{Galloping response and natural frequency}

 Now the oscillator equation Eq.\eqref{final_equation_motion} is considered from a power perspective. The shedding forces can be neglected because the net effect is negligible as system oscillates at natural frequency which is far from shedding frequency, for the cases that exhibit galloping. It is obvious that the forcing term on LHS of the equation is only dependent on transverse velocity($\dot{y}$) which is essentially the input power of the system. On the RHS, the mechanical damping or system damping is the only term that takes out power at any instant. This could be expressed as the product of the damping force and the velocity ($P_d$). The inertia and the stiffness terms governs the frequency of the system but the forces associated by those terms are conservative forces (i.e there is zero net energy in or out of the system when averaged over a period). Therefore the system is governed by the transverse velocity rather than the natural frequency.

 Using $U^*$ and $\zeta$ assumes that the system has a preferred frequency because the scale with the natural frequency of the system. The effect of fixing $\zeta$ and increasing $U^*$ actually decreases the damping constant for a fixed free-stream velocity. ($U^*=\frac{U}{f \times D}$, $\zeta= \frac{c}{2 m \omega_n}$ ). Both these effects leads to the multiple lines that are horizontally transpose when $\zeta$ is increased (Fig.\ref{fig:uncollapsed_data} (e) and (f)). Therefore the effect of $\zeta$ essentially scales up the damping coefficient for a fixed $U^*$.

 The data presented in Fig.\ref{fig:uncollapsed_data} for various damping ratios, $\zeta$, can be collapsed into a single line for a for a particular force characteristic curve (i.e $C_y$ vs $\theta$ curve). These collapsed curves were  obtained for the velocity amplitude  and power by plotting as functions of as a function of  the non dimensionalised  damping constant $\frac{c}{\rho\mathcal{A}U}$ 
(Fig \ref{fig:collpased_data} (a),(b),(c) and (d)).  This further emphasizes that the galloping system is not frequency dependent. It is possible to obtain a different similar power output at different values of \ustar when the damping constant, $\frac{c}{\rho\mathcal{A}U}$, is kept fixed. An example of this case, as shown in Fig.\ref{fig:time_hostory_velocity_same_power}, clearly show that this is a result of similar velocity amplitudes between cases if one were to disregard the high frequencies due to shedding. As mentioned earlier, it is the transverse velocity that determines the energy provided by the fluid forcing and the mechanical damping.

\input{../figure/collapsed_data}

\input{../figure/velocity_compare}

 Power could be expressed as the product of force and velocity. Therefore the transferred power form fluid-to-body could be expressed as $P_t=F_y\dot{y}$. Similarly the dissipated power due to the mechanical damping could be expressed as $P_d=(c\dot{y})\dot{y}$. The time average of these two quantities should be equal due to energy conservation, provided that the mechanical friction is neglected . Analysing the  time histories of $P_t $ and $P_d$ at key regions (Fig.\ref{fig:regions_1}) on the mean power vs $U^*$ provides a detailed explanation for the variation of the output power when the reduced velocity is increased. The key regions consists of region 1 where the $P_{mean}$ increases with \ustar, region 2 where $P_{mean}$ becomes maximum and region 3 where $P_{mean}$ decreases with \ustar. It has been established earlier that the damping factor is a function of $U^*$. Therefore it could be derived that $U^*$ is inversely proportional to damping coefficient. Hence the damping coefficient reduces when you move from region 1 to 3. Fig \ref{fig:lift_curves} (a) shows that $C_y$ and therefore instantaneous force rises until $4^0$ where it peaks and then falls and at around $6^0$ becomes negative. Maximum amount of power could be transferred within the peak region. At the region where the instantaneous force becomes negative it will be opposing the velocity $\dot{y}$. Data at $\zeta=0.1$, $m^*=40$ and \reynoldsnumber=165 (Fig.\ref{fig:power_time_histories}) are analysed as and example.

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{../FnP/sketch_1}
\caption{ Three key regions taken into account to analyse the time histories of power in a typical mean power vs. $U^*$ curve at $Re=165$. In region 1, high damping suppresses oscillation, hence the power output is low. In region 2, the damping is close to the optimum for power transfer. In region 3, the low damping means little energy is extracted from the fluid.}
\label{fig:regions_1}
\end{figure}

 At region 1 where $U^*=90$ the damping constant is high and a clear sinusoidal signal could be observed for both $P_d$ and $P_t$ in Fig.\ref{fig:power_time_histories} (a). Fig.\ref{fig:power_time_histories} (d) and (g) shows that $\theta$ is in line or in phase with $F_y$.  The velocity amplitude in this case is small and the equivalent incident angle within the range where the hydrodynamic force increases with the incident angle (i.e. $0<\theta \leq < 4^0$ in Fig.\ref{fig:lift_curves} (a)) Hence both $P_d$ and $P_t$ becomes sinusoidal. In this case, power output is limited by the low fluid forces present at low incident angles. In other words damping is significantly high and extracts a lot of power that the velocity amplitude could not grow where the forcing  is significant to produce high level of power.

  At region 3 ($U^*= 400$) `$c$' is low in comparison with region 1 and 2 which leads to a low mean power output. Fig.\ref{fig:power_time_histories} (c) shows that $P_t$ becomes negative over some portion of the cycle. This is because $\theta$  passes the point where both $\theta$ and $C_y$ (therefore $F_y$) are positive. The high velocity amplitude leads to the equivalent incident angle $\theta$, in this case to exceed the range where $C_y$ is positive (i.e. $0<\theta<6^0$ in Fig.\ref{cy ploynomial} (a)). In this portion of the cycle the hydrodynamic force actually opposes the direction of travel and power is transferred from the structure to the fluid during those times. From and energy perspective, the mechanical damping is not sufficient to remover the energy transferred from the fluid to the structure during other times of the cycle because $\frac{c}{\rho\mathcal{A}U}$ is substantially low. Therefore this excess energy is transferred back to the fluid as depicted by the negative region of $P_d$ in Fig.\ref{fig:power_time_histories} (c)

At region 2 ($U^*=165$). $P_t$ is not a pure sinusoidal signal. However, the  signal remains periodic. From the time history graph of $P_t$, two `peaks' are present in a single half cycle (Fig \ref{fig:power_time_histories} (b)). In this case, the velocity amplitude actually exceeds the equivalent incident angle where the hydrodynamic forces peaks (i.e. $\theta=4^0$ in \ref{cy ploynomial} (a)). The dips in $P_d$ between the two peaks approximately correspond to the time where the transverse velocity is higher than 0.07 and $F_y$ is decreasing with increasing transverse velocity. The mean power output is at its maximum. This is due to the fact that this region is a best compromise between region 1 and 3. The damping is substantially high to obtain a high power output while not too high to hinder the induced angle of attack entering the region where the forcing is high.

\input{../figure/time_histories}

\subsection{Effect of $m^*$}

 \input{../figure/m_star_collapsed}
 \input{../figure/mstar_parkinson}

The maximum mean power at different $m^*$ (Fig.\ref{fig:m_star_collapsed}(a)) was constant beyond $m^*=30$. However, at $m^* \leq 30$ an effect of $m^*$ it could be observed. Not only the peak value but overall reduction in the mean power could be observed when $m^*$ was reduced. However, when the sinusoidal forcing function in Eq.\ref{equationofmotion} which cater for the vortex shedding was disregarded, the reduction in power could not be observed Fig.\ref{fig:m_star_collapsed}(b). An influence of vortex shedding on power is present. This maybe due to the reduction in inertia as the mass ratio is decreased.

At \reynoldsnumber=22300 two major phenomena could be observed. The maximum of the mean power tend to increase  as the $m^*$ was decreased while keeping the $U^*$ constant (Fig\ref{fig:mstarcollapsed_parkinson} (a)). The same effect was observed when $U^*$ was increased keeping $m^*$ constant (\ref{fig:mstarcollapsed_parkinson} (b)). It should be noted that the influence of $U^*$ was observed only for low mass ratios.  The velocity time traces of example cases of both scenarios Fig.\ref{time_hostory_mstar_mass} and \ref{time_history_mstar_ustar} shows that essentially the same \hilight{phenomenon} occurring in both cases whereby the velocity signal tend to move towards a square from a sinusoidal signal. The displacement signal tend to become more triangular. the reason behind the increase in power while reducing mass ratio is due to the fact that the inertia of the system becomes low. Therefore the body accelerate faster and attains higher velocity quickly. Therefore the time period where the body sustains high velocity increases which will lead to an increase in time period where the body in influenced by the higher forcing, which leads to a higher power output. As for the case to where the mean power increases with increasing $U^*$ for a given low mass ratio, the exploration tend to be similar where the only difference being that the at high $U^*$ the period of oscillation tend to increase therefore the body sustains higher velocities for a longer time leading to a higher power output.

\input{../figure/time_history_msar_mass}

\input{../figure/time_history_msar_ustar}

\input{../figure/qss_fsi}

\subsection{Comparison with FSI simulations}
 Similar trends were captured for both displacement and velocity amplitudes between QSS and FSI simulations (Fig. \ref{fig:FSI_QSS_compare}(a) and \ref{fig:FSI_QSS_compare}(b)). Quantitatively a large discrepancy (average of $30\%$) could be observed between QSS and FSI data. Therefore the power also becomes significantly low in FSI data (Fig.\ref{fig:FSI_QSS_compare} (c)). However, the FSI data (Fig.\ref{fig:FSI_QSS_compare} (c)) were able to produce the main the rise and the fall of mean power when $U^*$ was increased. The reasoning behind this fact is that galloping is weak at \reynoldsnumber$=165$  and therefore fluid damping has a significant effect. It was reported by \cite{Barrero-Gil2009} that galloping only starts to occur ar Re $\geq 159$. As power is function of $(\dot{y})^2$ the error between QSS and FSI power becomes significantly large.

\section{Conclusion}

In this paper, the power transfer of a square body under aero elastic galloping is analysed by solving the quasi-steady state model equations through numerical integration. The power output of the system is not dependent on \ustar or natural frequency of the system, but dependent on the transverse velocity amplitude of the system. This is shown by the collapsed plots in terms of non-dimensionalised damping constant (using flow parameters). By analysing key regions of the power vs \ustar curve it could be concluded that in order to obtain an optimum power output, the damping constant ($\frac{c}{\rho\mathcal{A}U}$) should be high, but not excessive until it  to hinders the galloping from reaching induced angles of attack where the forcing is significant. The effect of mass ratio was could also be observed where at \reynoldsnumber=165 the mean power tend to decrease at $m^*<40$ which was found out to be an influence of vortex shedding. At \reynoldsnumber=22300 an opposite result could be observed where the mean power tend to increase with decreasing mass ratio as well as the mean power tend to increase with increasing $U^*$ at low mass ratios. The cause for this was found out to be the fact that as the mass ratio decreases, due to the lower inertia the velocity sinusoidal signal tend to move towards a square signal where it sustains high velocities for longer periods of time which leads to a higher mean power output.

 % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %