### 2013-11-22 research/paper/figure/coefficient-table.tex
Committing small changes to th mathematical model and some initial changes to the results section

\begin{table}[ht]

\begin{center}
\setlength{\unitlength}{\textwidth}

\begin{tabular}{c c c c c} % centered columns (4 columns)
\hline\hline %inserts double horizontal lines
\\[0.2ex]
Case & $a_1$ & $a_3$ & $a_5$ & $a_7$ \\ [0.8ex] % inserts table 
%heading
\hline 
\\[0.8ex]% inserts single horizontal line
Re=165 & 1.3 & 125.3 & 1825.73 & 8765.3 \\[0.8ex] % inserting body of the table
Re=22300 & 2.69 & 168 & 1670 & 59900 \\ [1ex] % [1ex] adds vertical space
\hline %inserts single line
\end{tabular}

\caption{Coefficient values used in the 7th order interpolation polynomial for high ($Re=22300$) and low ($Re=165$) Reynolds numbers. These data are used as input data to calculate the right-hand side of Eq. \ref{final_equation_motion} throughout this study.}

\label{table:cy-coefficients} % is used to refer this table in the text
\end{center}
\end{table}