### 2013-12-16 research/paper/nomenclature.tex

\section*{Nomenclature}
%\textbf{Nomenclature}

\begin{tabular}{ll}
$a_1,a_3,a_5,a_7$ & coefficients of the polynomial to determine $C_y$ \\ 
$A$ & displacement amplitude\\
$c$ & damping constant \\
$D$ & characteristic length (side length) of the cross section of the body \\
$f=\sqrt{k/m}/2\pi$ & natural frequency of the system \\
$F_y$ & instantaneous force normal to the flow \\ 
$F_0$& amplitude of the oscillatory force due to vortex shedding \\
$k$ & spring constant \\
$m$ & mass of the body \\
$m_a$ & added mass \\
$P_d$ & power dissipated due to mechanical damping  \\
$P_{mean}$ & mean power \\
$P_t$   & power transferred to the body by the fluid \\
$t$ & time \\
$U$ & freestream velocity \\
$U_i$ & Induced velocity \\
$y,\dot{y},\ddot{y}$ & transverse displacement, velocity and acceleration of the body \\
$\mathcal{A}=DL$ & frontal area of the body\\ 
$\rho$ & fluid density  \\
$\omega_n= 2 \pi f$ & natural angular frequency of the system  \\
$\omega_s$ & vortex shedding angular frequency \\
$\cstar=cD/mU$ & non-dimensionalised damping factor \\
$C_y=F_y/0.5\rho U^2DL$ & normal (lift) force coefficient \\
$m^*=m/\rho D^2L$ & mass ratio \\
$Re$ & Reynolds number  \\
$U^*=U/fD$ & reduced velocity  \\
$Y=y/D$ & non-dimensional transverse displacement \\
$\dot{Y}=m^*\dot{y}/a_1U$ & non-dimensional transverse velocity \\
$\ddot{Y}=m^{*2}D/a_1^2U^2$ & non-dimensional transverse acceleration \\
$\Gamma_1 = 4\pi^2m^{*2}/U^{*2}a_1^2$ & First dimensionless group arising from linearised, non-dimensionalised equation of motion\\
$\Gamma_2 = c^*m^*/a_1$ & Second dimensionless group arising from linearised, non-dimensionalised equation of motion\\
$\zeta= c/2 m \omega_n$ & damping ratio \\
$\theta= \tan^{-1}{(\dot{y}/U)}$ & instantaneous angle of incidence (angle of attack)\\
$\massstiff =  4\pi^2m^{*2}/U^{*2}$ & Combined mass-stiffness parameter\\
$\massdamp = c^*m^*$ & Combined mass-damping parameter\\
\end{tabular}